---
title: "alevinQC"
author: "Charlotte Soneson"
date: "`r Sys.Date()`"
package: "alevinQC"
output: 
    BiocStyle::html_document
bibliography: bibliography.bib
abstract: >
    `r Biocpkg("alevinQC")` reads output files from Alevin and 
    generates summary reports.
vignette: >
    %\VignetteIndexEntry{alevinQC}
    %\VignetteEncoding{UTF-8}  
    %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r v1, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = TRUE
)
```

# Introduction

The purpose of the `r Biocpkg("alevinQC")` package is to generate a summary QC
report based on the output of
[Alevin](https://salmon.readthedocs.io/en/latest/alevin.html)
[@Srivastava2018-fz]. The QC report can be generated as a html or pdf file, or
launched as a shiny application.

```{r v2, message=FALSE}
library(alevinQC)
```

# Assumed output directory structure

For more information about running Alevin, we refer to the
[documentation](https://salmon.readthedocs.io/en/latest/alevin.html). When
invoked, Alevin generates several output files in the specified output
directory. `r Biocpkg("alevinQC")` assumes that this structure is retained, and
will return an error if it doesn't  - thus, it is not recommended to move or
rename the output files from Alevin. `r Biocpkg("alevinQC")` assumes that the
following files (in the indicated structure) are available in the provided
`baseDir` (note that currently, in order to generate the full set of files,
Alevin must be invoked with the `--dumpFeatures` flag).

```{r v3, eval=FALSE, results='asis'}
baseDir
  |--alevin
  |    |--featureDump.txt
  |    |--filtered_cb_frequency.txt
  |    |--MappedUmi.txt
  |    |--quants_mat_cols.txt
  |    |--quants_mat_rows.txt
  |    |--quants_mat.gz
  |    |--raw_cb_frequency.txt
  |    |--whitelist.txt
  |--aux_info
  |    |--meta_info.json
  |--cmd_info.json
```

# Check that all required Alevin files are available

The report generation functions (see below) will check that all the required
files are available in the provided base directory. However, you can also call
the function `checkAlevinInputFiles()` to run the check manually. If all files
are available, this function returns nothing. If one or more files are missing,
the function will raise an error indicating the missing files.

```{r v4}
baseDir <- system.file("extdata/alevin_example", package = "alevinQC")
checkAlevinInputFiles(baseDir = baseDir)
```

# Generate QC report

The `alevinQCReport()` function generates the QC report from the Alevin output.
Depending on the file extension of the `outputFile` argument, and the value of
`outputFormat`, the function can generate either a html report or a pdf report.

```{r v5, message=FALSE, warning=FALSE}
outputDir <- tempdir()
alevinQCReport(baseDir = baseDir, sampleId = "testSample", 
               outputFile = "alevinReport.html", 
               outputFormat = "html_document",
               outputDir = outputDir, forceOverwrite = TRUE)
```

# Create shiny app

In addition to static reports, `r Biocpkg("alevinQC")` can also generate a shiny
application, containing the same information as the pdf and html reports.

```{r v7, message=FALSE}
app <- alevinQCShiny(baseDir = baseDir, sampleId = "testSample")
```

Once created, the app can be launched using the `runApp()` function from the 
`r CRANpkg("shiny")` package.

```{r v8, eval=FALSE}
shiny::runApp(app)
```

# Generate individual plots

The individual plots included in the QC reports can also be independently
generated. To do so, we must first read the Alevin output into an R object.

```{r v9}
alevin <- readAlevinQC(baseDir = baseDir)
```

The resulting list contains three entries:

- `cbTable`: a `data.frame` with various inferred characteristics of the
individual cell barcodes 
- `summaryTable`: a `data.frame` with summary information about the run 
- `versionTable`: a `data.frame` with information about the invokation of Alevin

```{r v10}
head(alevin$cbTable)
```

```{r v11}
knitr::kable(alevin$summaryTable)
```

```{r v12}
knitr::kable(alevin$versionTable)
```

The plots can now be generated using the dedicated plotting functions provided
with `r Biocpkg("alevinQC")` (see the help file for the respective function for
more information).

```{r v13, warning=FALSE}
plotAlevinKneeRaw(alevin$cbTable)
plotAlevinBarcodeCollapse(alevin$cbTable)
plotAlevinQuant(alevin$cbTable)
plotAlevinKneeNbrGenes(alevin$cbTable)
```

# Session info

```{r v14}
sessionInfo()
```

# References
